Efficient online Sequential Monte Carlo smoother for partially observed stochastic differential equations

### 2017-02-27 ParisEM.tex
modification of the lemma 1

In the case where $\qk$ is unknown, the filtering weights in \eqref{eq:importance:weights} then become:
\label{eq:random:weight}
\widehat{\omega}^{\ell}_k := \frac{\widehat{q}_k(\xi_{k-1}^{I^{\ell}_k},\xi^{\ell}_k;\zeta_k)g_k(\xi^{\ell}_k)}{\vartheta_k(\xi^{I^{\ell}_k}_{k-1}) p_k (\xi_{k-1}^{I^{\ell}_k},\xi^{\ell}_k)}\eqsp.
Therefore, to obtain a generalized random version of PaRIS, we only need to be able to sample from weight $\Lambda_k^N(i,\ell)$ in the case of SDE based HMM. If $\widehat{q}_k$ may be upper bounded almost surely by a constant $\hat{\sigma}_+$, this task is performed using the accept-reject mechanism given in Algorithm~\ref{alg:AR:unknownq}. Lemma~\ref{lem:AR:unbiased} states that replacing  $\qk$ by its unbiased approximation is a valid alternative to step~\eqref{it:PaRIS:sampleindex}.\\
\begin{lemma}
\label{lem:AR:unbiased}
Let $0\le k\le n-1$.  For all $1\le i \le N$ and all $1\le \ell \le \widetilde{N}$, the conditional probability distribution given $\mathcal{G}_{k+1}^N$ of the random variable $J_{k}^{i,\ell}$ produced by Algorithm~\ref{alg:AR:unknownq} is  $\Lambda_{k-1}^N(i,\cdot)$. 
\end{lemma}
\begin{proof}
See Appendix \ref{sec:append:proofs}.
\end{proof}
Note that the assumption that for all $1\le j\le N$, $\sup_y \hat{q}_k(\xi_k^j,y,\zeta_k)< \hat{\sigma}_+^k$ almost surely is actually sufficient to apply Algorithm \ref{alg:AR:unknownq}. This improvement keeps the linear complexity property of PaRIS algorithm and Algorithm \ref{alg:Ozaki:PaRIS}, and allows to increase significantly the acceptance rate of Algorithm \ref{alg:AR:unknownq}. %Moreover, the AR algorithm still stands if one can define 

@@

\begin{algorithm}[H]
\caption{Random weight accept-reject backward sampling}
\begin{algorithmic}
\FORALL{$i \in 1,\dots, N$}
\FORALL{$\ell \in 1,\dots, \widetilde N$} 
\STATE \textbf{\sc Sampling Step} \\
Sample independently $
\zeta_k$, $U\sim \mathcal{U}[0,1]$ and $J\in\{1,\ldots,N\}$ with probabilities proportional to $\{\widehat{\omega}_{k}^1,\dots,\widehat{\omega}_{k}^N\}$.
\IF{ $$U \leq \frac{\widehat{\qk}(\xi_{k}^J,\xi_{k+1}^i,\zeta_k)}{\hat{\sigma}_+},$$}
\STATE Set $J_k^{i,\ell} = J$.
\ELSE 
\STATE Return to \textbf{\sc Sampling Step}
\ENDIF
\ENDFOR
\ENDFOR
\end{algorithmic}
\label{alg:AR:unknownq}
\end{algorithm}

>>>>

In the case where $\qk$ is unknown, the filtering weights in \eqref{eq:importance:weights} then become:
\begin{equation}
\label{eq:random:weight}
\widehat{\omega}^{\ell}_k := \frac{\widehat{q}_k(\xi_{k-1}^{I^{\ell}_k},\xi^{\ell}_k;\zeta_k)g_k(\xi^{\ell}_k)}{\vartheta_k(\xi^{I^{\ell}_k}_{k-1}) p_k (\xi_{k-1}^{I^{\ell}_k},\xi^{\ell}_k)}\eqsp.
\end{equation}
Therefore, to obtain a generalized random version of PaRIS, we only need to be able to sample from weight $\Lambda_k^N(i,\ell)$ in the case of SDE based HMM.\\
We need to define the following assumption:\\
\[
\textbf{A1: }\forall~0\leq k\leq n,~\exists \hat{\sigma}_+ \text{ such that }\forall (x,y) \in\mathbb{R}^d\times  \mathbb{R}^d,~~\widehat{q}_k(x,y,\zeta_k)\leq \hat{\sigma_+}
\]
Then, the following lemma provides the missing step of a Generalized Random PaRIS (GRand Paris)
\begin{lemma}
\label{lem:AR:unbiased}
Let $0\le k\le n-1$ and $1\le i \le N$, and suppose $A_1$ holds may be upper bounded almost surely by a constant $\hat{\sigma}_+$. Let's define the random variable $J_k^i$ as follow:
\begin{algorithmic}
\STATE \textsc{Sampling step:} Sample independently $
\zeta_k$, $U\sim \mathcal{U}[0,1]$ and $J\in\{1,\ldots,N\}$ with probabilities proportional to $\{\widehat{\omega}_{k}^1,\dots,\widehat{\omega}_{k}^N\}$.
\IF{ $$U \leq \frac{\widehat{\qk}(\xi_{k}^J,\xi_{k+1}^i,\zeta_k)}{\hat{\sigma}_+},$$}
\STATE Set $J_k^{i,\ell} = J$.
\ELSE 
\STATE Return to \textbf{\sc Sampling Step}
\ENDIF
\end{algorithmic}
Then, the conditional probability distribution given $\mathcal{G}_{k+1}^N$ of $J_{k}^{i,\ell}$ is $\Lambda_{k-1}^N(i,\cdot)$. 
\end{lemma}
\begin{proof}
See Appendix \ref{sec:append:proofs}.
\end{proof}
Note that the assumption \textbf{A1} can be relaxed and replaced by one of the two following properties:
\begin{align*}
\textbf{A2: }\forall~0\leq k\leq n,~\exists \hat{\sigma}_+ \text{ such that }\forall (j,y) \in [[1,N]]\times  \mathbb{R}^d,~~\widehat{q}_k(\xi^j_k,y,\zeta_k)\leq \hat{\sigma_+}\\
\textbf{A3: }\forall~0\leq k\leq n,~\exists \hat{\sigma}_+ \text{ such that }\forall (j,y) \in [[1,N]]\times  \mathbb{R}^d,~~\widehat{q}_k(\xi^j_k,y,\zeta_k)\leq \hat{\sigma_+}
\end{align*}

 for all $1\le j\le N$, $\sup_y \hat{q}_k(\xi_k^j,y,\zeta_k)< \hat{\sigma}_+^k$ almost surely is actually sufficient to apply Algorithm \ref{alg:AR:unknownq}. This improvement keeps the linear complexity property of PaRIS algorithm and Algorithm \ref{alg:Ozaki:PaRIS}, and allows to increase significantly the acceptance rate of Algorithm \ref{alg:AR:unknownq}. %Moreover, the AR algorithm still stands if one can define 

@@
